one-sample t-test: compare a group of value to a certain value
two sample t-test: two groups compare with each other
paired t-test: e.g. test the blood pressure, test once and then treat the patients with medicine, after that test a second time.    same individuals and pick two samples

```{r}
library(readr)
kc<-read_csv("KungCensus.csv")
```

#create file with only females older than 18 years of age
```{r}
kaf <- subset(kc, sex=="woman" & age > 18)
summary(kaf$height)
```

```{r}
mean(kaf$height,na.rm=TRUE)
sd(kaf$height, na.rm = TRUE)
```

#One-sample t-test 
```{r}
t.test(kaf$height, mu=155)
t.test(kaf$height, mu=155, conf.level=0.99)
```
the size of the confidence interval is changed and the others remain the same.   the change of the CI is not large enough to extend to 14.

exercise
```{r}
t.test(kaf$weight, mu=40)
t.test(kaf$weight, mu=40, conf.level = 0.95)
t.test(kaf$weight, mu=40, conf.level = 0.99)
```
reject the null hypothesis.
Below is the same way of generate t-test
```{r}
km<-subset(kc, sex=="man")
kw<-subset(kc, sex=="woman")
t.test(km$weight, kw$weight)
```

#Two-sample t-test
```{r}
t.test(kc$weight ~ kc$sex)
t.test(kc$weight ~ kc$sex,conf.level=0.99)
```
reject the null hypothesis
#Paired t-test
#first, install and run package ISwR
```{r}
library(ISwR)
intake
t.test(intake$post, intake$pre, paired=T)
t.test(intake$pre, intake$post, paired=T)
```

```{r}
t.test(intake$post, intake$pre, paired=T,conf.level=0.99)
t.test(intake$post, intake$pre,conf.level = 0.95)
t.test(intake$post, intake$pre,conf.level = 0.99)
```

```{r}
kfm
```

```{r}
str(intake)
```

```{r}
kfm_man<-subset(kfm, sex=="boy")
kfm_woman<-subset(kfm, sex=="girl")
t.test(kfm_man$weight, kfm_woman$weight)
t.test(kfm_man$dl.milk, kfm_woman$dl.milk)

#t.test(kfm$weight~kfm$sex)

```

the data is from the same year. paired.
```{r}
Longevity<-read_csv("humanlongevity.csv")
t.test(Longevity$longevity.men,Longevity$longevity.women,paried=TRUE)
```







